Effects of rotation and magnetic braking on Li abundances for solar-type stars

### 2019-10-20 project/overleaf/paper.tex
Paper v3.3.0
Incorporando cambios tras revisión de Antonio y Adrián

% Abstract of the paper
\begin{abstract}
\textit{Goals.} The study of lithium (Li) surface abundance in the Sun and the influence of rotation and magnetic braking (MB) on its depletion during pre-main sequence (PMS) and main sequence (MS).
\newline\textit{Methods.} The effects of rotational mixing and of the rotational hydrostatic effects on Li abundances are studied by simulating several grids of PMS \& MS rotating and non-rotating models. The previous effects are combined with the additional impact of the MB (with magnetic field intensities ranging between 3.5 and 5.5 G) and the data obtained from the two set of simulations are confronted by comparing a number of parameters: Li abundances, angular velocity, effective temperature and size of convective zone.
\newline\textit{Results.} The surface Li abundance for the Sun like models at the end of the PMS and throughout the MS is decreased when rotational effects are included, i.e. the Li depletion rate for rotating models is higher than for non-rotating ones. However, this effect is attenuated when the braking effect produced by a magnetic field is present. This physical phenomena, generally neglected in star evolutionary models, impacts also the effective temperature ($T_{eff}$) of the star and therefore its location in the HR diagram. The impact of MB in Li depletion is sensitive to the magnetic field intensity: the higher it is, the lower the Li destruction. A direct link between the magnetic fields and the convective zone (CZ) size is observed: stronger magnetic fields produce shallower CZ's. Another important result of this study is that MB effect must be taken into consideration during PMS must if we aim to reproduce Li abundances in young clusters.
\end{abstract}

@@

The denominated standard models, models that include convection only as a mixing process and do not consider other transport options like diffusion or angular momentum loss (AML), have been mainly involved in the elaboration of these predictions \citep{Sestito2005}. Li is destroyed at a temperature $T_{Li} \approx 2.5 x 10^6\; K$, and is therefore directly destroyed in stellar envelopes when the temperature at the base of the convection zone (BCZ) reaches that value. The Sun in particular and the solar-type stars in general are characterized by having a CZ that covers much of the stellar radius during the PMS which causes that their lower limit to exceed $T_{Li}$ \citep{Iben1965}. This depletion stops during the approach to the zero-age MS (ZAMS) when the convection zone retreats and the temperature at BCZ is cooler than $T_{Li}$. In the standard stellar models only the mass and the initial chemical composition determine at what distance from the stellar surface $T_{Li}$ temperature is reached, therefore stars within clusters with similar mass are expected to reach the ZAMS with equal surface Li abundances. Furthermore, they should also show a very similar Li evolution until their approach to the terminal-age MS (TAMS). During this same period the convection mechanisms unleash a mixing process that homogenizes the chemical composition of the convective envelope, from its lower limit till the star surface. However, different abundances of Li have been observed for different star populations (\citet{Somers2014} and references therein).\par

From approximately the 1920s of the last century it has been accepted that rotation is capable of inducing the mixing of the elements present inside the stars. The studies of Eddington (1925) and Vogt (1925) had revealed the fact that the southern circulation must necessarily arise in those stars in rotation as a consequence of the von Zeipel paradox: a star cannot be simultaneously in thermal and hydrostatic equilibrium if its rotation speed depends exclusively on its radius. The solution to this depends on the existence of southern circulations. At the same time, this type of instability implies an impact on the convection mechanisms not considered by the standard models. Today it is known that in combination with this type of instability exists a catalogue of others that appear as a consequence of assuming models in rotation \citep{Maeder2003a}, particularly the differential rotation as a function of the radius of the star or the deviation of the spherical symmetry. Although initially these effects could be candidates to explain the observations, the fact is that the Sun is a slow rotator, meaning slow rotation that one for which the maximum centrifugal force is much smaller than the force of gravity. Furthermore, its deviation of the measured spherical symmetry is of the order of $10^{-5}$ and the time scale for the southern circulation is of the order of $10^{12}$ years, a time period well above the nuclear combustion time scale \citep{Pinsonneault1997}. In view of these findings, it seems evident that there are still open questions to be resolved on the subject of the processes that participate, directly or indirectly, in the mixing mechanisms that are not incorporated in the standard stellar models.\par

The impact of rotation both on PMS and Li depletion for solar-type stars has already extensively debated in the past (\citet{Pinsonneault1997}, \citet{Jeffries2004}, \citet{Somers2014}) and revised more recently on the basis of the availability of more accurate measures (\citet{Bouvier2016}). However, these previous studies focused mainly on the hydrostatic effects and didn't consider the influence of a coupling between the star magnetic field and its possible spin-down effect. It's thought that AML has a direct influence in the mixing processes and that transport of the momentum could be produced by a number of relevant mechanisms: mass loss, magnetic fields and gravity waves (aka, g-modes). Concerning the two latter, gravity waves \citep{Charbonnel2005} and magnetic fields \citep{Eggenberger2009} have the property of transmiting angular momentum (AM) much more effectively than inducing mixing \citep{Denissenkov2007}. As a consequence of this increment in the AM transport efficiency, the amount of differential rotation between the radiative and the convective zones of the star is reduced (a solid-body rotation is fostered), as are the induced rotational instabilities. Some relevant efforts have already been undertaken aimed to model their link with AM distribution (\citet{Eggenberger2008}, \citet{Fuller2014}). As result, the efficiency of these processes is most easily tested by their influence on AM evolution. Of particular interest for this work is the impact caused by the magnetic field of the star in the AML, more specifically the effect produced on the solar wind particles that are forced to rotate in sync with it. As a consequence of this coupling a torque is induced which slows down the star, i.e. the magnetic braking effect. These magnetic fields would originate in the limits between the radiative and convective zones of the star, in the so-called tachoclines, due to the rotational difference between both zones \citep{Guerrero2016}. \par

\subsection{Effects of rotation}
Rotation may affect the equations of stellar structure in four different ways according to \citet{Endal1976}:
%\renewcommand{\theenumi}{\arabic{enumi}.}
\begin{enumerate}
    \item The centrifugal force causes the local effective gravity ($g_{eff}$) associated with a point of the star not located on the axis of rotation to be reduced.
    \item In general, the centrifugal force will not be parallel to the gravitational force so that the equipotential layers will cease to be spherical.
    \item As the von Zeipel theorem states, the radiative flux in a uniformly rotating star is proportional to the $g_{eff}$, therefore the radiative flux is not constant on an equipotential surface.
    \item Rotation can suppress certain mechanisms that trigger convection, indirectly affecting mixing processes.
\end{enumerate}

Observational data support the fact that stars are born with different angular velocities even if they come from the same interstellar cloud. They also back up that they undergo a deceleration process throughout their evolution \citep{Skumanich1972} in a time period between 0.5 and 5 Gyr \citep{Somers2014a}. This range of different velocities implies that the mixing mechanisms induced by rotation vary their efficiency and effectiveness, and therefore have a direct influence on the abundances of the different chemical elements observed on the surface of stars. This fact, together with the differential degree of rotation to which the limits between the radiative and convective zones of the star are subjected, is specially interesting for our study since Li, as mentioned in the introduction, is destroyed at a low temperature. Therefore, variations in the intensity of the mixing processes can cause convective zones to extend to interior zones with temperatures higher than $T_{Li}$. The deceleration process to which a star is subjected throughout its evolution and the direct link that the rotation has with the mixing processes also offer a natural and consistent explanation for the fact that the destruction of Li is more accentuated or efficient in young stars than in old stars \citep{Sestito2005}. In a similar way, stellar rotation rates show a large dispersion at ZAMS \citep{Stauffer1984}, providing in this way a consistent argument for explaining the variance of Li abundances among stars which have a similar mass and were formed within the same cloud. Although this approach is not a novelty at all, it is no less true that in those studies based on it the effect caused by MB over the rotation on the PMS and MS is often overlooked. \par

Additionally, and according to the prevalent theory of rotational mixing, southern circulation and internal shears leads to a homogenization of the material depleted of Li found within the CZ, causing that the inner layers where there is a temperature higher than $T_{Li}$ and where Li is destroyed, been mixed with those cooler above them. These mechanism causes a positive feedback loop in which the cooler material sinks, is heated again and the remaining Li is destroyed, totally or partially, before proceeding to rise once more into cooler layers. \par

\subsection{Effects of magnetic fields}
The origin of magnetic fields is still unknown. The organized magnetic fields detected in some O stars \citep{Wade2010} might be fossil fields (see \citet{Dudorov2014} for details about the theory of fossil magnetic field), or fields produced through a dynamo mechanism \citep{Cantiello2009}. Besides that, the presence of magnetic fields with intensities of the order of kG \citep{Hussain2014} has been a sine qua non condition for dealing with certain data observed in T Tauri type stars in young, accreting PMS systems \citep{Johns-Krull2007}. One of this features is the long rotating periods  which is much longer than predicted from AM conservation. A key question is to understand how these young stars can accrete large amounts of material from their circumstellar disk with high specific angular momentum and at the same time they're able to keep slow rotation periods, between 7-10 days \citep{Hussain2014}. One plausible explanation for this behaviour considers the AML produced by the interaction between the magnetic field and the particles that form its circumstellar disk \citep{Zanni2012}. Once the accretion process stops or becomes less efficient then the star is able to spin up as it contracts towards the ZAMS.\par

The mass of stars plays a determining role in the existence or not of an external convective envelope. In turn, the presence of this convective region conditions the evolution of magnetic fields. Massive stars do not have external convective zones although they can expose a convective core. In these cases, the dominant type of energy transport in the outer zones of the star is radiative. On the other hand, the less massive stars such as the Sun, have convective covers and a radiative core. In both types of stars strong magnetic fields can be generated in the tachocline but with notable differences. For the former, these magnetic fields can survive for a long time but not reach the surface of the star. For the latter, these magnetic fields, although they are able to reach the surface of the star, decay quickly in time, in a few hundred years or even only in a few decades \citep{Chabrier2006}. It's therefore necessary a mechanism for generating magnetic field, e.g. a dynamo type, continued in time to explain its existence. The way in which this dynamo process is able to generate magnetic fields depends on different parameters and to this day this mechanism is still not fully accepted and is subject of controversy, e.g. \citet{Charbonneau2010}.\par

\subsubsection{Internal Magnetic Fields}
Our Sun is an active star as far as the magnetic field is concerned. According to the theoretical models this magnetic field is generated by a dynamo type effect in the tachocline, the intermediate zone between the radiative core and the CZ \citep{Aschwanden2014}, and has an intensity of approximately $B\approx10^5\, G$. Additionally, these models predict that because of the differential rotation existing between the core and CZ, the lines of the magnetic field that at the moment of their generation have a poloidal orientation will be affected by the differential rotation and forced to adopt a toroidal typology. This change of typology in the magnetic field, from poloidal to toroidal, is known as the $\omega$-effect. On the other hand, we have that the opposite case can also occur, that is to say, starting from a toroidal magnetic field to obtain a poloidal configuration. This would be possible on the one hand, thanks to the convective movements of material that are produced in the CZ of the star and on the other, to assume that this material behaves as a perfect conductor. This second condition allows to assume that the magnetic field lines are "frozen" in the conductive fluid and therefore must move in solidarity with it. Under these circumstances the magnetic field lines will be pushed to higher areas of the CZ as the material moves vertically and, due to the combined effect of stellar rotation plus Coriolis forces, the magnetic lines will be twisted and give rise to a poloidal magnetic field. This is what is known as the $\alpha$-effect.\par

A number of theoretical models have been proposed based on $\alpha$ and $\omega$ effects as an explanation to the generation of magnetic fields, among which is the so-called $\alpha\omega$ or also known as dynamo effect which was proposed by \citet{Spruit2002}. By means of this model it is possible to explain the change in polarity of the magnetic field that occurs in the Sun every 11 years, completing a full cycle every 22 years. It's also worth to mention that a theoretical debate on this topic is still ongoing \citep{Denissenkov2007} and with regards to MESA, an implementation according to the work of Spruit-Tayler has been coded \citep{Paxton2013}.\par

\subsubsection{Surface Magnetic Fields} \label{surf_mf}
The intensity of the magnetic field measured on the Sun's surface is about $B\approx1\, G$, twice the value of the Earth's one. Additionally its topology is varied and complex. As mentioned in the previous section, the dynamo model establishes that the magnetic field produced inside the star reaches the surface with the help of convective processes taken place in the CZ and gives rise to magnetically active regions on they reach the solar surface. Among them, we can find sunspots dominated by magnetic fields with an approximate intensity of $B\approx10^3\, G$ and coronal loops with field strengths of $B\approx10^2\, G$ at the photosphere, and $B\approx 10\, G$ in larger coronal heights \citep{Aschwanden2014}. Besides, rotating stars that expose a significant outer CZ can produce surface magnetic fields through a dynamo mechanism (e.g. \citet{Brandenburg2004}, \citet{Charbonneau2010} or \citet{Brun2017}). Whatever the origins of surface magnetic fields are, it's expected to couple to the wind mass-loss and, if this is strong enough, to produce MB (e.g. \citet{UdDoula2002}, \citet{Ud-Doula2007}, \citet{Ud-Doula2008} \citet{Meynet2010}) .\par

\begin{enumerate}
    \item Differential rotation: the AM transport is pushed by southern currents and shear instabilities.
    \item Solid body rotation: when the transport of AM is very efficient, the the solid body rotation is the prevailing mechanism maintained throughout the entire MS phase and the chemical elements are transported by the southern currents.
\end{enumerate}

The AML loss caused by MB depends directly on the amount of mass lost by the star due to the stellar winds. The estimated mass loss ratio for a solar-type star, approximately $10^{-13}M_{\odot} \; yr^{-1}$, and the resulting AML derived from that value (refer to \ref{mod_mb} for details) can be considered relatively modest as to decisively influence the evolution of the star. Furthermore, the combined effect of convective movements and differential rotation lead to the generation of magnetic fields. Those magnetic fields end up reaching the surface of the star \citep{Langer2012}, so that the AML can be increased in several orders of magnitude. That increment is as a consequence of the magnetic field that forces the stellar wind ionized particles to turn with the same angular speed, winds which extends a distance several times the radius of the star (see \citet{UdDoula2002}, \citet{Ud-Doula2007} and \citet{Ud-Doula2008} for more details).

\section{Method} \label{sec_2}
\begin{table}
	\centering
	\caption{Summary of adopted physics in MESA (based on \citet{Choi2016})}
	\label{tab:phy_mesa}
	\begin{tabular}{ll} 
		\hline
		Parameter & Adopted prescriptions and values\\
		\hline
		Solar Abundance & $X_{\odot}=0.7154, Y_{\odot}=0.2703, Z_{\odot}=0.0142$\\
		Equation of State & OPAL+SCVH+MacDonald+HELM+PC\\
		Opacity & OPAL Type I for log T $\geq$ 4 \\ & Ferguson for logT $\leq$ 4\\
		Reaction Rates & JINA REACLIB\\
		Boundary Conditions & ATLAS12; $\tau$=100 tables + photosphere\\
		Difussion & Track \isotope[1]{H}, \isotope[2]{He}, \isotope[7]{Li}, \isotope[7]{Be}\\
		Rotation & Differential rotation at PMS \& MS\\
		Convection & MLT + Ledoux, $\alpha_{MLT}$ = 1.82\\
		Overshoot & time-dependent, diffusive, \\ & $f_{ov,core}=0.0160$,\\ 
		& $f_{ov,sh}=0.0174$\\
		Semiconvection & $\alpha_{sc}=0.1$\\
		Thermohaline & $\alpha_{th}=666$\\
		Rotational Mixing & Include SH, ES, GSF, SSI \& DSI\\
		Magnetic Effects & Magnetic braking based on idealized \\ & monopole field\\
		Magnetic Field & B(G) variable between [3.5 - 5.5]\\
		Mass Loss & activated, $\Dot{M}_{max} = 10^{-3} \: M_{\sun} \: yr^{-1}$\\
		Angular Moment Loss & activated, $\Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{A}$\\
		\hline
	\end{tabular}
\end{table}

\subsection{Modelling rotation}
The rotation to which a star is subjected has profound implications for its inner structure and, therefore, for the way in which it evolves throughout its life. It is also responsible for inducing through hydrodynamic instabilities the mixing of the chemical elements that make it up. In order to be able to correctly model those induces mixing effects consistently with the observational data, it is necessary to analyze and understand more precisely the mechanisms that govern the distribution of AM (\citet{Pinsonneault1997}, \citet{Maeder2000}), since in this area the theoretical models also offer deficiencies to explain the observations \citep{Denissenkov2007}. \par

Angular velocity, or rather the way in which it evolves, also conditions the mechanism by which theoretical models solve the equations of stellar structure. It is evident that stars have a three-dimensional structure and for this reason the equations that model their evolution must take this fact into account. However, it has been proven sufficient to solve the equations in only one dimension, extrapolating the results to the other two, if it is fulfilled that the angular velocity is constant in regions subjected to the same pressure (isobaric). This is what is known as shellular approximation \citep{Meynet1997} and is based on calculating the impact of the centrifugal acceleration caused by the rotation in the stellar evolution equations \citep{Endal1976}. MESA uses this simplification to accelerate computational calculations. Additionally, as stated in \citet{Paxton2013}: "the transport of AM and chemical elements due to rotationally induced instabilities is implemented in a diffusion approximation \citet{Endal1978}. MESA calculates diffusion coefficients for five different rotationally induced mixing processes: dynamical shear instability, Solberg-H{\o}iland instability, secular shear instability, Eddington-Sweet circulation, and the Goldreich-Schubert-Fricke instability".\par

\subsection{Modelling magnetic braking} \label{mod_mb}
In the presence of mass-loss, AM is removed by the material in the stellar wind, resulting in a spin-down effect. This is quite important for example in massive stars, that are known to be rapidly rotating and have strong stellar winds. It has been also observed that strongly magnetic intermediate-mass stars typically have rotation rates much slower than other stars in their parent population \citep{Mathys2006}. In those stars, the presence of such magnetic field will interact with the mass loss and the Alfv\'{e}n radius ($R_{A}$) plays an important role in this interaction. The $R_{A}$ is defined as the point in which the magnetic field energy density $B^{2}/8\upi$ and the kinetic energy density $\frac{1}{2}\rho\nu^{2}$ of the stellar wind are balanced, that is, where the coefficient between both is equal to 1. In case that $R_{A}$ is greater than the star radius, then the wind flow will have to follow the magnetic field. As a consequence, the material will leave the stellar surface with a higher specific AM, as the co-rotation radius has increased and it will roughly correspond to the $R_{A}$. This effect is known as MB.\par

The energy ratio (eq. \ref{eq:wind_conf}) is denominated wind confinement magnetic parameter ($\eta_*$) and defines a characteristic parameter for the relative effectiveness of the magnetic fields in circumscribing and/or channeling the wind outflow \citep{UdDoula2002}. In turn, of the observable characteristics of the solar wind, we especially care about the parameters related to the amount of mass that is able to pull out of the outer layers of the star in a given time interval, and the speed that reaches the wind itself measured at a great distance from the star, the terminal velocity ($\nu_\infty$).\par

Each of these factors has a decisive influence on the future of the star and therefore in our routine implementation. On the one hand, stars with similar initial masses but different mass loss ratios will end up evolving very differently. On the other hand, the ionized particles carried by the solar wind not only contribute to the loss of mass but also to the loss of kinetic ($K_e$) energy that is deposited in the interstellar medium. Given a star with a spherically symmetric wind, the mass loss rate ($\Dot{M}$) and $K_e$ are characterized and related by the following expressions:

In general, a magnetically channeled outflow will have a complex stream geometry but for convenience, the expression (\ref{eq:mass_loss}) simply characterizes the wind strength in terms of a spherically symmetric mass-loss rate. $\nu$ can be characterized by the radial variation of outflow velocity in terms of the velocity law:
\begin{ceqn}
\begin{equation}
    \nu(r) = \nu_\infty (1-R_*/r)
\end{equation}
\end{ceqn}
where $\nu_\infty$ is the terminal wind velocity defined as the velocity that the wind, the outflowing matter, reaches at large distance from the central star, where it is not accelerated anymore by the wind driving force but its deceleration due to interaction with the interstellar medium (ISM) is negligible \citep{Niedzielski2002}.\par

The AML is directly influenced by the $R_A$ (or alternatively by the $\eta_*$), $\Omega$ and $\Dot{M}$. If the value of $\eta_*$ is large, the obtained AML will be also large and the star undergoes a considerable deceleration. In our proposal, the value of $\eta_*$ can be directly influenced by establishing in the different simulations the value of the magnetic field ($B$) on the surface or the rotation speed of the star ($\Omega$). With regards to $\Dot{M}$ and as reported in table (\ref{tab:phy_mesa}), for the set of simulations performed by MESA the empirical formula developed by Reimers for stars in the asymptotic giant branch (AGB) is used for calculating its value. For a solar-type star the $\Dot{M}$ during MS is relatively small, about  $10^{-13}M_{\odot} \, yr^{-1}$. Finally, $\Omega$ can be influenced in our models specifying the ratio $\Omega / \Omega_{crit}$ as another free parameter.

\subsection{Modelling angular velocity and angular momentum evolution} \label{mod_ang_vel_ang_mom}
As stated in the work of \citet{Paxton2015}, MESA makes its models evolve by default without taking into account the rotation but offers options to initialize them with a user defined angular speed. The initial value for $\Omega$ can be indicated as a function of the rotation speed of the star surface expressed in $km\,s^{-1}$ or using the critical rotation speed as a reference (option used in our case).  Once the model has been initialized with a specific $\Omega$, it is recalculated as the star evolves as a function of its mass, radius and AML that the simulator determines in each time step. \par

MESA assigns an $\Omega$ value to each of the cells that make up the star model. The number of cells does not remain fixed throughout the simulation but varies according to certain evolutionary parameters. The $\Omega$ for each cell $k$ ($\Omega_k$) is adjusted so that the resulting angular momentum is retained after adjusting the new mass of the cell $k$ ($m_k$), according to the loss or gain suffered by the solar wind, and the distance from that particular cell to the center of the star ($r_k$). After resolving the new interior structure of the star, an AM value is assigned to each cell $k$ ($J_k$). The evolution of the AM for each of these cells is performed after knowing the new radius of the star and, the mass and distance at which each cell is from the star center. Known $m_k$ and $r_k$ proceeds to calculate the AM of specific inertia ($I_k$), and from $I_k$ and using $J_k$ of the previous computational step, derives the new value of $\Omega_k$ by means of $J_k/I_k$ and thus manages to conserve the AM. It is only after these calculations that MESA derives the new value of $J_k$ (see \citet{Paxton2015} for more details). It's in this point when our MB routine plugs in and allows us to influence the internally calculated value by providing an additional contribution calculated by the user routine. Our implementation of the MB routine calculates the additional AML ($\Dot{J}_{k}$), derived from the external torque exerted by the magnetic field and distributes it through the different layers that make up the CZ according to the expression (\ref{eq:k_jdot}):\par

\section{Results} \label{sec_3}
\subsection{Angular velocity, Li abundance and effective temperature}
The destruction of Li occurs when a Li atom collides with a proton producing two He atoms, something that takes place during proton-proton type II reactions (P-P II) and at a temperature lower than that required for H fusion, as low as $2.5\, x\, 10^6\, K$. Those stars without a convective zone can still destroy Li by non-convective mixing processes which can take place in radiative regions and are driven by AML. Fast rotating ZAMS stars have suffered little AML and so would have the highest Li abundances. Slow rotators may have suffered low AML if they started with a less AM, or a lot if they were coupled to a circumstellar disk for an extended period of time \citep{Eggenberger2010}, and therefore may have a range of Li abundances. \par

@@

>>>>

% Abstract of the paper
\begin{abstract}
\textit{Goals.} The study of lithium (Li) surface abundance in the Sun and the influence of rotation and magnetic braking (MB) on its depletion during pre-main sequence (PMS) and main sequence (MS).
\newline\textit{Methods.} The effects of rotational mixing and of the rotational hydrostatic effects on Li abundances are studied by simulating several grids of PMS \& MS rotating and non-rotating models. The previous effects are combined with the additional impact of the MB (with magnetic field intensities ranging between 3.5 and 5.5 G) and the data obtained from the two set of simulations are confronted by comparing a number of parameters: Li abundances, angular velocity, effective temperature and size of convective zone.
\newline\textit{Results.} The surface Li abundance for the Sun like models at the end of the PMS and throughout the MS is decreased when rotational effects are included, i.e. the Li depletion rate for rotating models is higher than for non-rotating ones. However, this effect is attenuated when the braking effect produced by a magnetic field is present. This physical phenomena, generally neglected in star evolutionary models, impacts also the effective temperature ($T_{eff}$) of the star and therefore its location in the HR diagram. The impact of MB in Li depletion is sensitive to the magnetic field intensity: the higher it is, the lower the Li destruction. A direct link between the magnetic fields and the convective zone (CZ) size is observed: stronger magnetic fields produce shallower CZ's. Another important result of this study is that MB effect must be taken into consideration during PMS if we aim to reproduce Li abundances in young clusters.
\end{abstract}

@@

The denominated standard models, models that include convection only as a mixing process and do not consider other transport options like diffusion or angular momentum loss (AML), have been mainly involved in the elaboration of these predictions \citep{Sestito2005}. Li is destroyed at a temperature $T_{Li} \approx 2.5 x 10^6\; K$ when a Li atom collides with a proton producing two He atoms, something that takes place during proton-proton type II reactions (P-P II), and is therefore directly destroyed in stellar envelopes when the temperature at the base of the convection zone (BCZ) reaches that value. The Sun in particular and the solar-type stars in general are characterized by having a CZ that covers much of the stellar radius during the PMS which causes that their lower limit to exceed $T_{Li}$ \citep{Iben1965}. This depletion stops during the approach to the zero-age MS (ZAMS) when the convection zone retreats and the temperature at BCZ is cooler than $T_{Li}$. In the standard stellar models only the mass and the initial chemical composition determine at what distance from the stellar surface $T_{Li}$ temperature is reached, therefore stars within clusters with similar mass are expected to reach the ZAMS with equal surface Li abundances. Furthermore, they should also show a very similar Li evolution until their approach to the terminal-age MS (TAMS). During this same period the convection mechanisms unleash a mixing process that homogenizes the chemical composition of the convective envelope, from its lower limit till the star surface. However, different abundances of Li have been observed for different star populations \citep[][and references therein]{Somers2014}.\par

@@

The impact of rotation both on PMS and Li depletion for solar-type stars has already extensively debated in the past \citep{Pinsonneault1997,Jeffries2004,Somers2014} and revised more recently on the basis of the availability of more accurate measures \citep{Bouvier2016}. However, these previous studies focused mainly on the hydrostatic effects and didn't consider the influence of a coupling between the star magnetic field and its possible spin-down effect. It's thought that AML has a direct influence in the mixing processes and that transport of the momentum could be produced by a number of relevant mechanisms: mass loss, magnetic fields and gravity waves (g-modes). Concerning the two latter, gravity waves \citep{Charbonnel2005} and magnetic fields \citep{Eggenberger2009} have the property of transmiting angular momentum (AM) much more effectively than inducing mixing \citep{Denissenkov2007}. As a consequence of this increment in the AM transport efficiency, the amount of differential rotation between the radiative and the convective zones of the star is reduced (a solid-body rotation is fostered), as are the induced rotational instabilities. Of particular interest for this work is the impact caused by the magnetic field of the star in the AML, more specifically the effect produced on the solar wind particles that are forced to rotate in sync with it. As a consequence of this coupling a torque is induced which slows down the star, i.e. the magnetic braking effect. These magnetic fields would originate in the limits between the radiative and convective zones of the star, in the so-called tachoclines, due to the rotational difference between both zones \citep{Guerrero2016}. \par

Today, standard models are not able to replicate in a satisfactory way the observed values of Li abundance on star surfaces with the predictions that these models yield. This lack of agreement makes us think that certain physical mechanisms that influence the destruction of Li are either being modelled improperly or simply not being taken into account, e.g. the magnetic braking. In view of these findings, it seems evident that there are still open questions to be resolved on the subject of the processes that participate, directly or indirectly, in the mixing mechanisms that are not incorporated in the standard stellar models.\par

According to the prevalent theory of rotational mixing, southern circulation and internal shears leads to a homogenization of the material depleted of Li found within the CZ, causing that the inner layers where there is a temperature higher than $T_{Li}$ and where Li is destroyed, been mixed with those cooler above them. These mechanism causes a positive feedback loop in which the cooler material sinks, is heated again and the remaining Li is destroyed, totally or partially, before proceeding to rise once more into cooler layers. \par

\subsection{Effects of magnetic fields}
The origin of magnetic fields is still unknown. The organized magnetic fields detected in some O stars \citep{Wade2010} might be fossil fields \citep[see][for details about the theory of fossil magnetic field]{Dudorov2014}, or fields produced through a dynamo mechanism \citep{Cantiello2009}. Besides that, the presence of magnetic fields with intensities of the order of kG \citep{Hussain2014} has been a sine qua non condition for dealing with certain data observed in T Tauri type stars in young, accreting PMS systems \citep{Johns-Krull2007}.\par

The mass of stars plays a determining role in the existence or not of an external convective envelope. In turn, the presence of this convective region conditions the evolution of magnetic fields. Massive stars do not have external convective zones although they can expose a convective core. On the other hand, the less massive stars such as the Sun, have convective covers and a radiative core. In both types of stars strong magnetic fields can be generated in the tachocline but with notable differences (see \citep{Chabrier2006} \citet{Charbonneau2010}\ for details).\par

\subsubsection{Internal Magnetic Fields}
Our Sun is an active star as far as the magnetic field is concerned. According to the theoretical models this magnetic field is generated by a dynamo type effect in the tachocline, the intermediate zone between the radiative core and the CZ \citep{Aschwanden2014}, and has an intensity of approximately $B\approx10^5\, G$. Additionally, these models predict that because of the differential rotation existing between the core and CZ, the lines of the magnetic field that at the moment of their generation have a poloidal orientation will be affected by the differential rotation and forced to adopt a toroidal typology (the $\omega$-effect). On the other hand, we have that the opposite case can also occur, that is to say, starting from a toroidal magnetic field to obtain a poloidal configuration. This would be possible on the one hand, thanks to the convective movements of material that are produced in the CZ of the star and on the other, to assume that this material behaves as a perfect conductor. This second condition allows to assume that the magnetic field lines are "frozen" in the conductive fluid and therefore must move in solidarity with it ($\alpha$-effect).\par

\subsubsection{Surface Magnetic Fields} \label{surf_mf}
The intensity of the magnetic field measured on the Sun's surface is about $B\approx1\, G$, twice the value of the Earth's one. Additionally its topology is varied and complex. The dynamo model establishes that the magnetic field produced inside the star reaches the surface with the help of convective processes taken place in the CZ and gives rise to magnetically active regions on they reach the solar surface. Besides, rotating stars that expose a significant outer CZ can produce surface magnetic fields through a dynamo mechanism \citep[e.g.][]{Brandenburg2004,Charbonneau2010,Brun2017}. Whatever the origins of surface magnetic fields are, it's expected to couple to the wind mass-loss and, if this is strong enough, to produce MB \citep[e.g.][]{UdDoula2002,Ud-Doula2007,Ud-Doula2008,Meynet2010}.\par

\begin{enumerate}
    \item Differential rotation: the AM transport is pushed by southern currents and shear instabilities.
    \item Solid body rotation: when the transport of AM is very efficient, the solid body rotation is the prevailing mechanism maintained throughout the entire MS phase and the chemical elements are transported by the southern currents.
\end{enumerate}

The AML caused by MB depends directly on the amount of mass lost by the star due to the stellar winds. The estimated mass loss ratio for a solar-type star, approximately $10^{-13}M_{\odot} \; yr^{-1}$, and the resulting AML derived from that value (refer to \ref{mod_mb} for details) can be considered relatively modest as to decisively influence the evolution of the star. Furthermore, the combined effect of convective movements and differential rotation lead to the generation of magnetic fields. Those magnetic fields end up reaching the surface of the star \citep{Langer2012}, so that the AML can be increased in several orders of magnitude. That increment is as a consequence of the magnetic field that forces the stellar wind ionized particles to turn with the same angular speed, winds which extends a distance several times the radius of the star \citep[see][for more details]{UdDoula2002,Ud-Doula2007,Ud-Doula2008}.

\section{Method} \label{sec_2}
\begin{table}
	\centering
	\caption{Summary of adopted physics in MESA \citep[based on][]{Choi2016}.}
	\label{tab:phy_mesa}
	\begin{tabular}{ll} 
		\hline
		Parameter & Adopted prescriptions and values\\
		\hline
		Solar Abundance & $X_{\odot}=0.7154, Y_{\odot}=0.2703, Z_{\odot}=0.0142$\\
		Equation of State & OPAL+SCVH+MacDonald+HELM+PC\\
		Opacity & OPAL Type I for log T $\geq$ 4 \\ & Ferguson for logT $\leq$ 4\\
		Reaction Rates & JINA REACLIB\\
		Boundary Conditions & ATLAS12; $\tau$=100 tables + photosphere\\
		Difussion & Track \isotope[1]{H}, \isotope[2]{He}, \isotope[7]{Li}, \isotope[7]{Be}\\
		Rotation & Differential rotation at PMS \& MS\\
		Convection & MLT + Ledoux, $\alpha_{MLT}$ = 1.82\\
		Overshoot & time-dependent, diffusive, \\ & $f_{ov,core}=0.0160$,\\ 
		& $f_{ov,sh}=0.0174$\\
		Semiconvection & $\alpha_{sc}=0.1$\\
		Thermohaline & $\alpha_{th}=666$\\
		Rotational Mixing & Include SH, ES, GSF, SSI \& DSI\\
		Magnetic Effects & Magnetic braking based on idealized \\ & monopole field\\
		Magnetic Field & B(G) variable between [3.5 - 5.5]\\
		Mass Loss & activated, $\Dot{M}_{max} = 10^{-3} \: M_{\sun} \: yr^{-1}$\\
		Angular Moment Loss & activated, $\Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{A}$\\
		\hline
	\end{tabular}
\end{table}

\subsection{Modelling rotation}
The rotation to which a star is subjected has profound implications for its inner structure and, therefore, for the way in which it evolves throughout its life. It is also responsible for inducing through hydrodynamic instabilities the mixing of the chemical elements that make it up. In order to be able to correctly model those induces mixing effects consistently with the observational data, it is necessary to analyze and understand more precisely the mechanisms that govern the distribution of AM \citep{Pinsonneault1997,Maeder2000}, since in this area the theoretical models also offer deficiencies to explain the observations \citep{Denissenkov2007}. \par

Stars have a three-dimensional structure and the equations that model their evolution must take this fact into account. However, it has been proven sufficient to solve the equations in only one dimension, extrapolating the results to the other two (shellular approximation \citep{Meynet1997}) MESA uses this simplification to accelerate computational calculations \citet{Paxton2013}.\par

\subsection{Modelling magnetic braking} \label{mod_mb}
In this section, we describe a physically sound yet computationally simple model extension to Modules for Experiments in Stellar Astrophysics star evolution code (MESA; \citeauthor{Paxton2011} \citeyear{Paxton2011}, \citeyear{Paxton2013}, \citeyear{Paxton2015}, \citeyear{Paxton2018}, \citeyear{Paxton2019}) for the calculation of the AML as a result of the torque applied by a magnetically-coupled stellar winds. In its simplest form, the implementation of the model requires the specification of both the surface magnetic field strength ($B$) and the initial rotation rate ($\Omega$) at star surface. The models include already rotation during the PMS as there evidences that advocate for a strong established relationship between Li destruction and rotation on that phase (\citeauthor{Bouvier2016} \citeyear{Bouvier2016}, \citeyear{Bouvier2018}), and the numerical simulations trace the rotational history and A(Li) of a $1\, M_{\sun}$ star for a variety of initial values for $B$ and $\Omega$.\par
